---
title: "Exam 1"
author: "Lilly Clark"
date: "03-06-20"
output: 
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA, 
                      message = FALSE, warning = FALSE)
```

## Packages

```{r packages}
library(tidyverse)
library(purrr)
library(dplyr)
library(lubridate)
library(stringr)
```

## Data

```{r load-data}
tdf <- readRDS(file = "data/tdf_2019.rds")
```

### Task 1

```{r harvest-stage-info}
compile_stage <- function(n){
  stage_info <- data.frame(tdf[[n]]$stage) %>%
    select(description:single_event)
  
  people <- lapply(tdf[[n]]$stage$competitors, data.frame) %>%
    bind_rows() %>%
    mutate(description = (str_c("Stage ", n)))
  
  team_info <- lapply(tdf[[n]]$stage$teams, data.frame) %>%
    bind_rows() %>%
    rename(team.id = id)
  
  people_teams <- full_join(people, team_info, by = "team.id")
  
  whole_stage <- full_join(people_teams, stage_info, by = "description") 

  return (whole_stage)
}

messy_tdf <- lapply(1:21, compile_stage) %>%
  rbind_list()
```

```{r rename-relevant-variables}
messy_tdf <- messy_tdf %>%
  select(stage = description, rider_name = name.x, rider_nat = nationality.x, team_name = team.name, team_nat = team.nationality, dep_city = departure_city, arr_city = arrival_city, classification, distance, start_date = scheduled, time = result.time, time_rank = result.time_ranking, sprint_pts = result.sprint, sprint_rank = result.sprint_ranking, climb_pts = result.climber, climb_rank = result.climber_ranking, young_rider_time = result.young_rider, young_rider_rank = result.young_rider_ranking)
```

```{r correct-variable-types}
tdf_clean <- messy_tdf %>%
  mutate(stage = as.integer(str_extract(stage, "\\d.*")),
         distance = as.numeric(str_replace(distance, ",", ".")), 
         start_date = date(str_extract(start_date, "\\d{4}-\\d{2}-\\d{2}")),
         time_rank = as.integer(time_rank), 
         sprint_pts = as.numeric(sprint_pts),
         sprint_rank = as.integer(sprint_rank),
         climb_pts = as.numeric(climb_pts),
         climb_rank = as.integer(climb_rank),
         young_rider_rank = as.integer(young_rider_rank)
         )
```


```{r save-data-to-file}
saveRDS(tdf_clean, "results/tdf_clean.rds")
```

### Task 2

#### Part 1

```{r}
tdf_clean %>%
  filter(!is.na(young_rider_rank)) #seems to be an error, time is greater for rank 27 than 28 in stage 1
```

```{r convert-seconds-function}
convert_seconds <- function(time) {
  
  time <- str_remove(time, "\\+")
  
  minutes <- as.numeric(str_extract(time, "\\d{2}"))
  
  seconds <- str_extract(time, ":\\d{2}\\.") %>%
    str_extract(., "\\d{2}") %>%
    as.numeric()
  
  milliseconds <- str_extract(time, "\\.\\d{2}") %>%
    str_extract(., "\\d{2}") %>%
    as.numeric()
  
  time_in_seconds <- 60 * minutes + seconds + milliseconds / 100
  
  return (time_in_seconds)
}

tdf_clean <- tdf_clean %>%
  mutate(time = convert_seconds(time),
         young_rider_time = convert_seconds(young_rider_time))
```


compile_winners <- function(time_var, rank_var) {
  
  winning_times <- tdf_clean %>%
  filter(deparse(substitute(rank_var)) == 1) %>%
  select(stage, substitute(time_var))

  two <- c(stage = 2, deparse(substitute(time_var)) = NA)
  nineteen <- c(stage = 19, deparse(substitute(time_var)) = NA)

  winning_times <- rbind_list(winning_times, two, nineteen) %>%
    arrange(stage)
  
  return (winning_times)
}

compile_winners(tdf_clean$time, tdf_clean$time_rank)

```{r}
winning_times <- tdf_clean %>%
filter(time_rank == 1) %>%
select(stage, time)

two <- c(stage = 2, time = NA)
nineteen <- c(stage = 19, time = NA)

winning_times <- rbind_list(winning_times, two, nineteen) %>%
  arrange(stage)

yr_winning_times <- tdf_clean %>%
filter(young_rider_rank == 1) %>%
select(stage, young_rider_time)

yr_two <- c(stage = 2, young_rider_time = NA)
yr_nineteen <- c(stage = 19, young_rider_time = NA)

yr_winning_times <- rbind_list(yr_winning_times, yr_two, yr_nineteen) %>%
  arrange(stage)
```

```{r}
tdf_clean <- tdf_clean %>%
  mutate(time = case_when(
    stage == 1 & time_rank == 1 ~ time,
    stage == 1 & time_rank != 1 ~ winning_times[[1, 2]] + time,
    stage == 2 & time_rank == 1 ~ time,
    stage == 2 & time_rank != 1 ~ winning_times[[2, 2]] + time,
    stage == 3 & time_rank == 1 ~ time,
    stage == 3 & time_rank != 1 ~ winning_times[[3, 2]] + time,
    stage == 4 & time_rank == 1 ~ time,
    stage == 4 & time_rank != 1 ~ winning_times[[4, 2]] + time,
    stage == 5 & time_rank == 1 ~ time,
    stage == 5 & time_rank != 1 ~ winning_times[[5, 2]] + time,
    stage == 6 & time_rank == 1 ~ time,
    stage == 6 & time_rank != 1 ~ winning_times[[6, 2]] + time,
    stage == 7 & time_rank == 1 ~ time,
    stage == 7 & time_rank != 1 ~ winning_times[[7, 2]] + time,
    stage == 8 & time_rank == 1 ~ time,
    stage == 8 & time_rank != 1 ~ winning_times[[8, 2]] + time,
    stage == 9 & time_rank == 1 ~ time,
    stage == 9 & time_rank != 1 ~ winning_times[[9, 2]] + time,
    stage == 10 & time_rank == 1 ~ time,
    stage == 10 & time_rank != 1 ~ winning_times[[10, 2]] + time,
    stage == 11 & time_rank == 1 ~ time,
    stage == 11 & time_rank != 1 ~ winning_times[[11, 2]] + time,
    stage == 12 & time_rank == 1 ~ time,
    stage == 12 & time_rank != 1 ~ winning_times[[12, 2]] + time,
    stage == 13 & time_rank == 1 ~ time,
    stage == 13 & time_rank != 1 ~ winning_times[[13, 2]] + time,
    stage == 14 & time_rank == 1 ~ time,
    stage == 14 & time_rank != 1 ~ winning_times[[14, 2]] + time,
    stage == 15 & time_rank == 1 ~ time,
    stage == 15 & time_rank != 1 ~ winning_times[[15, 2]] + time,
    stage == 16 & time_rank == 1 ~ time,
    stage == 16 & time_rank != 1 ~ winning_times[[16, 2]] + time,
    stage == 17 & time_rank == 1 ~ time,
    stage == 17 & time_rank != 1 ~ winning_times[[17, 2]] + time,
    stage == 18 & time_rank == 1 ~ time,
    stage == 18 & time_rank != 1 ~ winning_times[[18, 2]] + time,
    stage == 19 & time_rank == 1 ~ time,
    stage == 19 & time_rank != 1 ~ winning_times[[19, 2]] + time,
    stage == 20 & time_rank == 1 ~ time,
    stage == 20 & time_rank != 1 ~ winning_times[[20, 2]] + time,
    stage == 21 & time_rank == 1 ~ time,
    stage == 21 & time_rank != 1 ~ winning_times[[21, 2]] + time,
    TRUE ~ time
  ))

tdf_clean <- tdf_clean %>%
  mutate(young_rider_time = case_when(
    stage == 1 & young_rider_rank == 1 ~ young_rider_time,
    stage == 1 & young_rider_rank != 1 ~ yr_winning_times[[1, 2]] + young_rider_time,
    stage == 2 & young_rider_rank == 1 ~ young_rider_time,
    stage == 2 & young_rider_rank != 1 ~ yr_winning_times[[2, 2]] + young_rider_time,
    stage == 3 & young_rider_rank == 1 ~ young_rider_time,
    stage == 3 & young_rider_rank != 1 ~ yr_winning_times[[3, 2]] + young_rider_time,
    stage == 4 & young_rider_rank == 1 ~ young_rider_time,
    stage == 4 & young_rider_rank != 1 ~ yr_winning_times[[4, 2]] + young_rider_time,
    stage == 5 & young_rider_rank == 1 ~ young_rider_time,
    stage == 5 & young_rider_rank != 1 ~ yr_winning_times[[5, 2]] + young_rider_time,
    stage == 6 & young_rider_rank == 1 ~ young_rider_time,
    stage == 6 & young_rider_rank != 1 ~ yr_winning_times[[6, 2]] + young_rider_time,
    stage == 7 & young_rider_rank == 1 ~ young_rider_time,
    stage == 7 & young_rider_rank != 1 ~ yr_winning_times[[7, 2]] + young_rider_time,
    stage == 8 & young_rider_rank == 1 ~ young_rider_time,
    stage == 8 & young_rider_rank != 1 ~ yr_winning_times[[8, 2]] + young_rider_time,
    stage == 9 & young_rider_rank == 1 ~ young_rider_time,
    stage == 9 & young_rider_rank != 1 ~ yr_winning_times[[9, 2]] + young_rider_time,
    stage == 10 & young_rider_rank == 1 ~ young_rider_time,
    stage == 10 & young_rider_rank != 1 ~ yr_winning_times[[10, 2]] + young_rider_time,
    stage == 11 & young_rider_rank == 1 ~ young_rider_time,
    stage == 11 & young_rider_rank != 1 ~ yr_winning_times[[11, 2]] + young_rider_time,
    stage == 12 & young_rider_rank == 1 ~ young_rider_time,
    stage == 12 & young_rider_rank != 1 ~ yr_winning_times[[12, 2]] + young_rider_time,
    stage == 13 & young_rider_rank == 1 ~ young_rider_time,
    stage == 13 & young_rider_rank != 1 ~ yr_winning_times[[13, 2]] + young_rider_time,
    stage == 14 & young_rider_rank == 1 ~ young_rider_time,
    stage == 14 & young_rider_rank != 1 ~ yr_winning_times[[14, 2]] + young_rider_time,
    stage == 15 & young_rider_rank == 1 ~ young_rider_time,
    stage == 15 & young_rider_rank != 1 ~ yr_winning_times[[15, 2]] + young_rider_time,
    stage == 16 & young_rider_rank == 1 ~ young_rider_time,
    stage == 16 & young_rider_rank != 1 ~ yr_winning_times[[16, 2]] + young_rider_time,
    stage == 17 & young_rider_rank == 1 ~ young_rider_time,
    stage == 17 & young_rider_rank != 1 ~ yr_winning_times[[17, 2]] + young_rider_time,
    stage == 18 & young_rider_rank == 1 ~ young_rider_time,
    stage == 18 & young_rider_rank != 1 ~ yr_winning_times[[18, 2]] + young_rider_time,
    stage == 19 & young_rider_rank == 1 ~ young_rider_time,
    stage == 19 & young_rider_rank != 1 ~ yr_winning_times[[19, 2]] + young_rider_time,
    stage == 20 & young_rider_rank == 1 ~ young_rider_time,
    stage == 20 & young_rider_rank != 1 ~ yr_winning_times[[20, 2]] + young_rider_time,
    stage == 21 & young_rider_rank == 1 ~ young_rider_time,
    stage == 21 & young_rider_rank != 1 ~ yr_winning_times[[21, 2]] + young_rider_time,
    TRUE ~ young_rider_time
  ))
```


minutes_and_seconds <- function(time) {
  min <- time %/% 60
  sec <- time %% 60
  return (str_c(min, ":", sec))
  as.POSIXlt(., tz = "", format = "%M:%S\\.%S")
}

```{r return-seconds-to-time}
tdf_clean <- tdf_clean %>%
  mutate(time = round(seconds_to_period(time), 2),
         young_rider_time = round(seconds_to_period(young_rider_time), 2))
```

#### Part 2

```{r}
mountains <- tdf_clean %>%
  select(stage, rider_name, climb_pts) %>%
  mutate(climb_pts = case_when(
    is.na(climb_pts) ~ 0,
    TRUE ~ climb_pts
  ),
  stage = str_c("stage_", stage))

mountains <- pivot_wider(mountains, id_cols = rider_name, names_from = stage, values_from = climb_pts)
```

```{r}
mountains <- mountains %>%
  mutate(total_climb_pts = select(., stage_1:stage_21) %>% 
           rowSums()) %>%
  arrange(desc(total_climb_pts)) %>%
  head(30)

rank <- c(1:30)

mountain_king <- cbind(mountains, rank)
mountain_king <- mountain_king[,c(1, 24, 2:23)]
```

```{r}
mountain_king %>%
  mutate(rank = as.numeric(rank))
```

 Only include the top 30 climbers sorted by total_climb_points in your final data frame. You may decide how to account for ties. Explain your choice.

### References 

Converting to decimal hours:
https://www.r-bloggers.com/using-dates-and-times-in-r/

deparse thang
https://stackoverflow.com/questions/14577412/how-to-convert-variable-object-name-into-string

row sums
https://stackoverflow.com/questions/29006056/efficiently-sum-across-multiple-columns-in-r

